---
title: ''
author: "Jenny Smith"
date: "February 2, 2018"
output: html_document
---


```{r setup}
library(knitr)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE, fig.align='center')
knitr::opts_knit$set(root.dir = '/fh/fast/meshinchi_s/workingDir/TARGET/AML_TARGET/RNA/miRNAseq/analysis/2018.01.29_mir130a_Expression_Outcome/')
```


```{r message = FALSE, warning=FALSE}
library(stringr)
library(magrittr)
library(ggplot2)
library(dplyr)
library(tibble)
library(tidyr)
getwd()
```


#Read in the Clinical Data

```{r}
CDE.1031 <- read.csv("~/reference_mapping-files/older_versions/TARGET_AML_1031_CDE_cleaned_1.31.2018.csv", stringsAsFactors = FALSE, row.names = 1)

head(CDE.1031[,1:5])
```


#Read in the miRNA Expression Matrix

```{r}
RPM <- read.csv("/fh/fast/meshinchi_s/workingDir/TARGET/AML_TARGET/RNA/miRNAseq/level3/matureMiRNA/2017July_BCCA_1031_miRBase20_Illumina_data/TARGET_AML_AAML1031_expn_matrix_mimat_miRNA_RPM.csv",
                stringsAsFactors = FALSE, row.names = 1)

head(RPM[,1:5])
# dim(RPM) #2576 by 1,138
```

```{r}
t <- str_split_fixed(colnames(RPM), "\\.", n=3)

table(t[,2]) #all are diagnostic samples 
```

```{r}
colnames(RPM) <- t[,1] #set colnames  and remove the Kasumis/MV411
RPM <- RPM[,grep("BCCA", colnames(RPM), invert = TRUE)]
```


#read in the miRNA annotations

```{r}
mirAnno <- read.csv("/fh/fast/meshinchi_s/workingDir/TARGET/Reference_Data/miRBase_v21/hsa_gff3_IDMap.csv", 
                    stringsAsFactors = FALSE)
head(mirAnno)
```



#waterfall plot with mir130

```{r}
mir130 <- RPM %>%
  rownames_to_column("mir") %>%
  filter(grepl("mir.130a", mir)) %>%
  gather(USI,value, -mir) %>%
  mutate(log2.RPM=log2(value+1),
         group=ifelse(grepl("^BM|^RO", USI), "NBM", "AML")) %>%
  left_join(., select(CDE.1031, USI,t.8.21_Cleaned,inv.16_Cleaned,MLL_Cleaned,Abnormal.Karyotype, osi,yrsos,efsi,yrsefs), by="USI") %>%
  
  mutate(CBF.Group=ifelse(t.8.21_Cleaned == "Positive", "t.8.21", ifelse(inv.16_Cleaned == "Positive", "inv.16", ifelse(MLL_Cleaned == "Positive", "MLL", ifelse(Abnormal.Karyotype == "Normal", "NK", group))))) %>%
  mutate(CBF.Group=ifelse(is.na(CBF.Group), group, CBF.Group)) %>%
  mutate(CBF.Group=factor(CBF.Group, levels=c("NBM", "AML", "NK", "inv.16","t.8.21", "MLL")))

head(mir130)
# dim(mir130)
```

In the future, will create my ordering function. Need to reorder USI so that the factor levels order BM and AML seperately, and BM and AML ordered by increasing RPM. 


```{r}
write.csv(mir130,"TARGET_AML_1031_miR130a_CytoGroups_Original.csv", row.names = FALSE)
```

```{r}
table(mir130$mir)
```

```{r}
library(ggpubr)
source("~/scripts/RNAseq_Analysis/DifferentialExpn_PathwayAnalysis/ggplot_Themes_Function.r")
```


```{r}
mimat <- c("MIMAT0000425", "MIMAT0004593")
waterfalls <- list()

for (i in 1:2){
  p <- ggbarplot(subset(mir130, grepl(mimat[i], mir130$mir)), x="USI", y="value", 
          fill = "group", 
          color = "group", 
          palette = c("firebrick", "black"),
          sort.by.groups = TRUE,
          sort.val = "asc",
          xlab="Patient", 
          ylab="RPM",
          title = paste("miR-130",mimat[i], "Expression in AAML1031"))
  
  waterfalls[[i]] <- p
}

```


```{r}
# tiff("mir130a_MIMAT0000425_waterfallplot.tiff", height = 5, width = 8, units="in", res = 600)
waterfalls[[1]] + theme_JS + theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), panel.border = element_blank())
# dev.off()

# tiff("mir130a_MIMAT0004593_waterfallplot.tiff", height = 5, width = 8, units="in", res = 600)
waterfalls[[2]] + theme_JS + theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), panel.border = element_blank()) + ylim(0,1500)
# dev.off()
```


```{r}
hist <- ggplot(mir130, aes(x=log2.RPM, fill=group)) +
  geom_histogram(binwidth = 0.5) +
  scale_fill_manual(values=c("NBM"="black", "AML"="firebrick")) +
  facet_wrap(facets = "mir", ncol=2)
  # scale_x_continuous(breaks = seq(0,10,by=1)) + 
  # geom_vline(xintercept = c(0.25, 5), color="grey30") +
  # theme_numX 

# tiff("")
hist
```



```{r}
bplot <- ggplot(mir130, aes(x=CBF.Group, y=log2.RPM, fill=CBF.Group)) +
  geom_boxplot() 

bplot <- bplot + 
  facet_wrap(facets = "mir", ncol=2, strip.position = "bottom") +
  theme(strip.background = element_rect(color = "deepskyblue1", fill = "deepskyblue1", size = 0.5), 
        strip.text = element_text(size=18)) + 
  theme_JS +
  theme(axis.title.x = element_blank(), legend.text = element_text(size=18), legend.title = element_blank())


# tiff("mir130a_Cytogenetic_Groups_boxplot.tiff", height = 8, width = 12, res=600, units="in")
bplot
# dev.off()
```

```{r}
groups <- subset(mir130,grepl("MIMAT0004593", mir))

table(groups$Abnormal.Karyotype)

table(groups$CBF.Group)
# write.csv(as.data.frame(table(groups$CBF.Group)), "Boxplot_Figure_legend.csv", row.names = FALSE)
```

```{r}

```


#Heatmap with Mir130a 

```{r}

```



#Survival analysis with mir130a

```{r}
source("~/scripts/survival_analysis/Survplot_Functions_2017.03.31.r")
```


```{r}
expn_groups <- function(numericVector,USI,type="median"){
  
  if (type == "median"){
    m <- median(numericVector)
    g <- ifelse(numericVector >= m, "high", "low")
  }else if (type=="quantile"){
    q <- quantile(numericVector)
    
    if (q[1] == q[2]){
      #add USIs to the ordered numeric vector
      numline <- numericVector %>% set_names(USI) %>% .[order(.)] #ordered from lowest to highest value
      qts <- quantile(1:length(numline))
      q1 <- numline[1:qts[2]-1] #identify USIs in q1
      
      #update the groups so that q1s are included
      g <- ifelse(numericVector < q[2], "q1", ifelse(numericVector >= q[2] & numericVector < q[3], "q2", ifelse(numericVector >= q[3] & numericVector < q[4], "q3", "q4"))) %>%
        set_names(USI)
      g <- ifelse(names(g) %in% names(q1), "q1", g)
    }else{
     g <- ifelse(numericVector < q[2], "q1", ifelse(numericVector >= q[2] & numericVector < q[3], "q2", ifelse(numericVector >= q[3] & numericVector < q[4], "q3", "q4"))) 
    }
  }
  
  return(g)
}
```


```{r}
mir130.withOS <- mir130 %>%
  filter(!is.na(osi)) %>%
  # inner_join(., select(CDE.1031.withOS, USI,osi,yrsos,efsi,yrsefs), by="USI") %>%
  mutate(value=log2(value+1)) %>%
  group_by(mir) %>%
  mutate(High.Low.Median=expn_groups(value,USI),
         High.Low.Quartile=expn_groups(value,USI, type="quantile"))


head(mir130.withOS)
```


#Continous Variable

```{r}
library(survival)
```

```{r}
df <- mir130.withOS %>%
  # group_by(mir) %>%
  filter(grepl("MIMAT0000425", mir)) 

KM <- survival::Surv(time = df$yrsos, event =df$osi)
coxMod <- coxph(KM ~ df$value)
coxPHtest <- cox.zph(coxMod, transform = "km")

summary(coxMod)
coxPHtest
```


There is not a relationship with outcome. 


#Discretized Expression Groups

```{r message=FALSE}
KM.med <- KM.plots(mir130.withOS, 
                   groupBy = "mir", 
                   type="OS",
                   covariate = "High.Low.Median",
                   cohort = "1031", 
                   cc = c("low"="blue", "high"="red"))
KM.med
# saveMultiPlots(KM.med, w=11)
```


```{r fig.width=11}
KM.med$OS
KM.med$EFS
```


```{r}
KM.qt <- KM.plots(mir130.withOS, 
                  groupBy = "mir", 
                  type="OS",
                  covariate = "High.Low.Quartile", 
                  cohort = "1031", 
                  cc=c("q1"="blue", "q2"="dodgerblue", "q3"="red1","q4"="red4" ))

KM.qt

# saveMultiPlots(KM.qt,w=11)
```


```{r fig.width=11}
KM.qt$OS
KM.qt$EFS
```


#Session Information 

```{r}
sessionInfo()
```

