---
title: 'miR-130a in Corebinding Factor AML'
author: "Jenny Smith"
date: "February 2, 2018"
output: html_document
---


```{r setup}
library(knitr)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE, fig.align='center')
knitr::opts_knit$set(root.dir = '/fh/fast/meshinchi_s/workingDir/TARGET/AML_TARGET/RNA/miRNAseq/analysis/2018.01.29_mir130a_Expression_Outcome/')
options(stringsAsFactors = FALSE)
```

```{r message = FALSE, warning=FALSE}
library(stringr)
library(magrittr)
library(ggplot2)
library(gridExtra)
library(dplyr)
library(tibble)
library(tidyr)
getwd()
```


#Read in the Clinical Data

```{r}
CDE.1031 <- read.csv("~/reference_mapping-files/AAML1031_Target_CDEs_with_HiAR_and_PrimaryCyto_withFusionCalls_10.11.2018.csv")

CDE.1031 <- CDE.1031 %>%
  filter(!is.na(USI)) %>%
  set_rownames(.$USI)

head(CDE.1031[,1:5])
dim(CDE.1031)
```

```{r}
merged <- read.csv("~/reference_mapping-files/TARGET_AML_0531_1031_merged_CDEs_10.29.18.csv")


merged <- merged %>%
  select(USI=TARGET.USI.1, everything()) %>%
  filter(!is.na(USI)) %>%
  set_rownames(.$USI)

head(merged[,1:5])
dim(merged) #2153  176
```


#Read in the miRNA Expression Matrix

```{r}
RPM <- read.csv("~/RNA_seq_Analysis/0000.00.03_Expression_Matrices/TARGET_AML_AAML1031_MIMAT_mirbasev20_miRNA_RPM.csv")

head(RPM[,1:5])
dim(RPM) #2576 1138
```



#read in the miRNA annotations

```{r}
mirAnno <- read.csv("/fh/fast/meshinchi_s/workingDir/TARGET/Reference_Data/miRBase_v21/hsa_gff3_IDMap.csv", 
                    stringsAsFactors = FALSE)
head(mirAnno)
```



#Melt MiR-130a  and Define Cohort

```{r}
mir130 <- RPM %>%
  filter(grepl( "mir.130a",mir)) %>%
  gather(TARGET.Barcode,RPM, -mir) %>%
  filter(grepl("^BM|^RO|^P[A-Z]", TARGET.Barcode)) %>% #ensure only AML and NBM (no cell lines)
  
  mutate(log2.RPM=log2(RPM+1),
         group=ifelse(grepl("^BM|^RO", TARGET.Barcode), "NBM", "AML"),
         USI=str_split_fixed(TARGET.Barcode, "\\.", n=3)[,1], 
         mimat=str_split_fixed(mir,"\\.", 2)[,2]) %>%
  
  left_join(., mirAnno, by=c("mimat"="MIMAT.ID")) %>%
  left_join(., select(merged, USI,Protocol, ISCN,Primary.Cytogenetic.Code,
                      Cytogenetic.Category.1, Cytogenetic.Category.2, Rare.Fusions,
                      Overall.Survival.Time.in.Days, Event.Free.Survival.Time.in.Days,
                      OS.ID,Event.ID), by="USI") %>%
  
  mutate_at(vars(Cytogenetic.Category.1),
            funs(case_when(
              group=="NBM" ~ "NBM",
              #fix MLLs who are pos. by karyotype but classified as "other" in primary cytogenetic code.
              USI %in% c("PAUUXA", "PAVGIX", "PAVNUJ", "PAVSFB", "PAVSLI") ~ "MLL",
              
              #fix one t(8,21) who was listed as normal due to a cryptic insersion.
              grepl("cryp ins(21)(8;21)", ISCN, fixed = TRUE)  ~ "t.8.21.",  
              
              grepl("Other",.) ~ "OtherAML",
              is.na(.) ~ "OtherAML", 
              grepl("Unknown", .) ~ "OtherAML",
              TRUE ~ .))) %>%
  mutate(Cytogenetic.Category.1=factor(Cytogenetic.Category.1, levels=c("NBM" ,
                                                                        "t.8.21.",
                                                                        "inv.16.",
                                                                        "MLL",
                                                                        "Normal",
                                                                        "OtherAML"))) %>%
 mutate_at(vars(Cytogenetic.Category.2, 
                 Rare.Fusions, 
                 Protocol),
            funs(case_when(
              group=="NBM" ~ "NBM",
              is.na(.) ~ "OtherAML", 
              grepl("Unknown", .) ~ "OtherAML",
              TRUE ~ .))) %>%

  mutate(CBF.AML=case_when(
    grepl("t.8.21.|inv.16.", Cytogenetic.Category.1) ~ "CBF.AML", 
    grepl("NBM", Cytogenetic.Category.1) ~ "NBM",
    TRUE ~ "OtherAML"))
  

# head(mir130)
# dim(mir130) 2270   18
```



In the future, will create my ordering function. Need to reorder USI so that the factor levels order BM and AML seperately, and BM and AML ordered by increasing RPM. 

```{r}
# table(mir130$group[mir130$mir=="hsa-mir-130a.MIMAT0000425"], useNA = "always")
# table(mir130$Protocol[mir130$mir=="hsa-mir-130a.MIMAT0000425"], useNA = "always")
table(mir130$Cytogenetic.Category.1[mir130$mir=="hsa-mir-130a.MIMAT0000425"], useNA = "always")
# table(mir130$Cytogenetic.Category.2[mir130$mir=="hsa-mir-130a.MIMAT0000425"], useNA = "always")
table(mir130$Rare.Fusions[mir130$mir=="hsa-mir-130a.MIMAT0000425"], useNA = "always")
table(mir130$CBF.AML[mir130$mir=="hsa-mir-130a.MIMAT0000425"], useNA = "always")
```


#Define Expression Groups (High vs Low)

```{r}
source("~/scripts/survival_analysis/Survplot_Functions_2018.10.24.r")
```


```{r}
expn_groups <- function(numericVector,USI,type="median"){
  
  if (type == "median"){
    m <- median(numericVector)
    g <- ifelse(numericVector > m, "high", "low")
    names(g) <- USI
    
  }else if (type=="quantile"){
    q <- quantile(numericVector)
    
    #if first 2 quartiles are all zeros    
    if(q[1]==q[2] & q[1]==q[3]){
      #add USIs to the ordered numeric vector
      numline <- numericVector %>% set_names(USI) %>% .[order(.)] #ordered from lowest to highest value
      qts <- round(quantile(1:length(numline)), digits = 0)
      
      q1 <- numline[(1:qts[2]-1)] #identify USIs in q1
      q2 <- numline[qts[2]:(qts[3]-1)] #identify USIs in q2
      
      #update the groups so that q1s are included
      grp <- ifelse(numericVector < q[2], "q1", 
                  ifelse(numericVector >= q[2] & numericVector < q[3], "q2", 
                         ifelse(numericVector >= q[3] & numericVector < q[4], "q3", "q4"))) %>%
        set_names(USI)
      
      g <- ifelse(names(grp) %in% names(q1), "NoExpn", grp)
      g <- ifelse(names(grp) %in% names(q2), "NoExpn", g)
      
    } else if (q[1] == q[2]){ #if first quartiles are all zeros 
      #add USIs to the ordered numeric vector
      numline <- numericVector %>% set_names(USI) %>% .[order(.)] #ordered from lowest to highest value
      qts <- quantile(1:length(numline))
      q1 <- numline[1:qts[2]-1] #identify USIs in q1
      
      #update the groups so that q1s are included
      g <- ifelse(numericVector < q[2], "q1", 
                  ifelse(numericVector >= q[2] & numericVector < q[3], "q2", 
                         ifelse(numericVector >= q[3] & numericVector < q[4], "q3", "q4"))) %>%
        set_names(USI)
      
      g <- ifelse(names(g) %in% names(q1), "NoExpn", g)
      
      
    }else{
     g <- ifelse(numericVector < q[2], "q1",
                 ifelse(numericVector >= q[2] & numericVector < q[3], "q2", 
                        ifelse(numericVector >= q[3] & numericVector < q[4], "q3", "q4"))) 
    }
  }
  
  return(g)
}
```

```{r}
mir130.expn <- mir130 %>%
  group_by(miR,Cytogenetic.Category.1) %>%
  
  mutate(High.Low.Median=expn_groups(log2.RPM,USI),
         High.Low.Quartile=expn_groups(log2.RPM,USI, type="quantile")) %>%
  
  mutate(High.25Per=ifelse(High.Low.Quartile=="q4", "q4", "q123"))


# head(mir130.expn)
# dim(mir130.expn)
```

```{r}
cbf.mirExpn <- mir130 %>%
  filter(grepl("CBF", CBF.AML )) %>%
  group_by(miR) %>%
  
  mutate(High.Low.Median=expn_groups(log2.RPM,USI),
         High.Low.Quartile=expn_groups(log2.RPM,USI, type="quantile")) %>%
   mutate(High.25Per=ifelse(High.Low.Quartile=="q4", "q4", "q123"))


# head(cbf.mirExpn)
# dim(cbf.mirExpn) #502

# write.csv(cbf.mirExpn,"TARGET_AML_1031_miR130a_Expression.csv")
```


#Boxplots of Expression

```{r fig.width=8, fig.height=5}
# png("TARGET_AML_1031_mir130a_Expn_Barplots_12.11.18.png", height = 5, width = 8, units="in", res=200)
ggplot(mir130.expn, aes(y=log2.RPM, x=Cytogenetic.Category.1, fill=Cytogenetic.Category.1)) +
  geom_boxplot() +
  # geom_histogram(binwidth = 0.1) +
  facet_wrap(~ miR,scales = "fixed", ncol = 2) +
  theme_bw() + 
  theme(strip.background = element_blank(),
        strip.text = element_text(size=16), 
        axis.title = element_text(size=20),
        axis.text = element_text(size=20, color="black"), 
        axis.text.x = element_text(angle = 45, hjust=1, vjust=1, color="black"))

# dev.off()
```

```{r fig.height=4, fig.width=5}
p1 <-  ggplot(cbf.mirExpn, aes(x=Cytogenetic.Category.1, y=log2.RPM, fill=Cytogenetic.Category.1))+
  geom_boxplot() +
  labs(title="Both miR-130a Species") +
  theme_bw()

p2 <- ggplot(cbf.mirExpn, aes(x=Cytogenetic.Category.1, y=log2.RPM, fill=Cytogenetic.Category.1))+
  geom_boxplot() +
  facet_wrap(~ miR) +
  theme_bw() +
  theme(strip.background = element_blank()) 

grid.arrange(p1,p2)
```





#Survival Analysis in Cytogenetic Groups

```{r}
mir130.expn.AML <- filter(mir130.expn, Cytogenetic.Category.1 != "NBM") %>%
  droplevels(.)
```

```{r}
mir130.expn.AML %>%
  group_by(mir) %>%
  summarise(N.Surv.Dat=sum(!is.na(Overall.Survival.Time.in.Days)))
```

##Continous Variable

```{r}
library(survival)
```

```{r}
df <- mir130.expn.AML %>%
  group_by(miR, Cytogenetic.Category.1) %>%
  do(KM = survival::Surv(time = .$Overall.Survival.Time.in.Days, event = .$OS.ID), 
     coxMod = coxph(Surv(time = .$Overall.Survival.Time.in.Days, event = .$OS.ID) ~ .$log2.RPM),
     coxPHtest = cox.zph(coxph(Surv(time = .$Overall.Survival.Time.in.Days, event = .$OS.ID) ~ .$log2.RPM),
                         transform = "km")) %>%
  arrange(Cytogenetic.Category.1)

df
```

```{r}
names(df$coxMod) <- paste0(df$miR, " in ", df$Cytogenetic.Category.1)
lapply(df$coxMod, summary)
```




##Median Groups 

```{r}
expnGroups <-  KM.plots(mir130.expn.AML,
                   "miR", "Cytogenetic.Category.1", #grouping variables
                   type="OS",
                   covariate = "High.Low.Median",
                   cohort = "0531", 
                   riskTable = FALSE,
                   cc = c("low"="blue", "high"="red")) %>%
  arrange(Cytogenetic.Category.1)

expnGroups
```

```{r fig.height=20, fig.width=11}
# png("TARGET_AML_1031_miR.130a_OS_by_Cyto.png", height = 20, width=11,units = "in", res=200)
grid.arrange(grobs = expnGroups$OS, ncol=2, nrow=5)
# dev.off()
```


```{r fig.height=20, fig.width=11}
 # png("TARGET_AML_1031_miR.130a_EFS_by_Cyto.png", height = 20, width=11,units = "in", res=200)
grid.arrange(grobs = expnGroups$EFS, ncol=2, nrow=5)
# dev.off()
```

```{r}
cbf.expnGroups <- KM.plots(cbf.mirExpn,
                    "miR", #grouping varaibles
                   type="OS",
                   covariate = "High.Low.Median",
                   cohort = "0531", 
                   riskTable = FALSE,
                   cc = c("low"="blue", "high"="red")) 
cbf.expnGroups
```

```{r fig.height=8, fig.width=6}
# png("TARGET_AML_1031_miR.130a_OS_in_CBF-AML.png", height =8, width=6,units = "in", res=200)
grid.arrange(grobs = cbf.expnGroups$OS, nrow=2)
# dev.off()
```

```{r fig.height=8, fig.width=6}
# png("TARGET_AML_1031_miR.130a_EFS_in_CBF-AML.png", height =8, width=6,units = "in", res=200)
grid.arrange(grobs = cbf.expnGroups$EFS,  nrow=2)
# dev.off()
```


##by quartile groups

c("q1"="blue", "q2"="dodgerblue", "q3"="red1","q4"="red4","NoExpn"="darkgrey")

```{r}
expnGroups.qt <-  KM.plots(mir130.expn.AML,
                   "miR", "Cytogenetic.Category.1", #grouping variables
                   type="OS",
                   covariate = "High.Low.Quartile",
                   cohort = "0531", 
                   riskTable = FALSE,
                   cc = NULL) %>%
  arrange(Cytogenetic.Category.1)

expnGroups.qt
```


```{r fig.height=20, fig.width=11}
# png("TARGET_AML_1031_miR.130a_OS_by_Cyto_quartiles.png", height = 20, width=11,units = "in", res=200)
grid.arrange(grobs = expnGroups.qt$OS, ncol=2, nrow=5)
# dev.off()
```


```{r fig.height=20, fig.width=11}
 # png("TARGET_AML_1031_miR.130a_EFS_by_Cyto_quartiles.png", height = 20, width=11,units = "in", res=200)
grid.arrange(grobs = expnGroups.qt$EFS, ncol=2, nrow=5)
# dev.off()
```

```{r}
cbf.expnGroups.qt <- KM.plots(cbf.mirExpn, 
                    "miR", "CBF.AML", #grouping varaibles
                   type="OS",
                   covariate = "High.Low.Quartile",
                   cohort = "0531", 
                   riskTable = FALSE,
                   cc = NULL) %>%
  arrange(CBF.AML)

cbf.expnGroups.qt
```

```{r fig.height=8, fig.width=11}
# png("TARGET_AML_1031_miR.130a_OS_in_CBF-AML_quartiles.png", height =8, width=11,units = "in", res=200)
grid.arrange(grobs = cbf.expnGroups.qt$OS, ncol=2, nrow=2)
# dev.off()
```

```{r fig.height=8, fig.width=11}
# png("TARGET_AML_1031_miR.130a_EFS_in_CBF-AML_quartiles.png", height =8, width=11,units = "in", res=200)
grid.arrange(grobs = cbf.expnGroups.qt$EFS, ncol=2, nrow=2)
# dev.off()
```


##by 4th Quartile vs All Others groups

c("q1"="blue", "q2"="dodgerblue", "q3"="red1","q4"="red4","NoExpn"="darkgrey")

```{r}
expnGroups.25 <-  KM.plots(mir130.expn.AML,
                   "miR", "Cytogenetic.Category.1", #grouping variables
                   type="OS",
                   covariate = "High.25Per",
                   cohort = "0531", 
                   riskTable = FALSE,
                   cc = NULL) %>%
  arrange(Cytogenetic.Category.1)

expnGroups.25
```


```{r fig.height=20, fig.width=11}
# png("TARGET_AML_1031_miR.130a_OS_by_Cyto_quartiles.png", height = 20, width=11,units = "in", res=200)
grid.arrange(grobs = expnGroups.25$OS, ncol=2, nrow=5)
# dev.off()
```

```{r fig.height=20, fig.width=11}
 # png("TARGET_AML_1031_miR.130a_EFS_by_Cyto.png", height = 20, width=11,units = "in", res=200)
grid.arrange(grobs = expnGroups.25$EFS, ncol=2, nrow=5)
# dev.off()
```

```{r}
cbf.expnGroups.25 <- KM.plots(cbf.mirExpn, 
                    "miR", "CBF.AML", #grouping varaibles
                   type="OS",
                   covariate = "High.25Per",
                   cohort = "0531", 
                   riskTable = FALSE,
                   cc = NULL) %>%
  arrange(CBF.AML)

cbf.expnGroups.25
```

```{r fig.height=8, fig.width=11}
# png("TARGET_AML_1031_miR.130a_OS_in_CBF-AML_quartiles.png", height =8, width=11,units = "in", res=200)
grid.arrange(grobs = cbf.expnGroups.25$OS, ncol=2, nrow=2)
# dev.off()
```

```{r fig.height=8, fig.width=11}
# png("TARGET_AML_1031_miR.130a_EFS_in_CBF-AML_quartiles.png", height =8, width=11,units = "in", res=200)
grid.arrange(grobs = cbf.expnGroups.25$EFS, ncol=2, nrow=2)
# dev.off()
```



#Session Information 

```{r}
sessionInfo()
```

